---
title: "Data Viz Webinar"
output: html_notebook
---

# Set WD
```{r}
setwd("C:/Users/Ashley Rainbow/Desktop/Data Fellowship/data-fellowship/Data Visualisation/Data Viz Webinar")
```

# Load packages
```{r}
library(tidyverse)
```

# Read in EPL data
```{r}
players <- read.csv("transformed_epl_data - transformed_epl_data.csv")
```

# Check the data using 'head'
```{r}
head(players)
```

# 1. Transform the data if needed

# 2. Start doing exploratory data analysis
## qplot which stands for quick plot
```{r}
qplot(data = players, x = players$market_value, y = players$page_views, geom = c("point", "smooth"))
```

```{r}
qplot(data = players, x = players$market_value, y = players$page_views, colour = players$position_cat)
```

```{r}
qplot(players$age, geom = "histogram", binwidth = 1)
```

# 3. Create a scatter plot using ggplot2 package
```{r}
ggplot(data = players, aes(x = players$market_value, y = players$page_views, label = players$name))+geom_point()+geom_text(size = 2)
```

## Continue to add layers
`vjust justifies vertically, aligned from bottom`
`check_overlap automatically reduces overlap by eliminating some labels`
`nudge moves the text up a bit to avoid it overlapping with the point`
`angle tilts the text to a set angle`
```{r}
ggplot(data = players, aes(x = players$market_value, y = players$page_views,
                           label = players$name, col = players$position_cat))+geom_point()+geom_text(check_overlap = T, vjust = "bottom", nudge_y = 200, angle = 30, size = 3)+labs(x = "Market Value", y = "Page Views", title = "Market Value vs Page Views of EPL players")
```

```{r}
player_score <- as.numeric(players$fpl_points/players$fpl_value)
players <- cbind(players, player_score)
players <- players[order(players$player_score)]
View(players)
```

```{r}
barplot(players$player_score)
```

