---
title: "Regression Practice"
output: html_notebook
---

#Overview
Regression is used to find relationships between and independant variable and a dependant variable.  
This can be shown as: *y = B^~0~ + B^~1~X*

#Brief
Creating a model to predict income.
#Set up
loading pROC, car, caTools, ggplot2, plotly and tidyverse
```{r}
#pROC is used for visualisatio, smoothing and comparing receiver operating characteristic.
library(pROC)
#car calculates type-II or type-III analysis-of-variance tables for model objects
library(car)
#caTools contains several basic utility functions
library(caTools)
library(ggplot2)
library(plotly)
library(tidyverse)
```

#Import the 'Prestige' dataset.
```{r}
prestige <- read.csv("Prestige.csv")
head(prestige)
```
```{r}
str(prestige)
```

#Visualise the Prestige data set using a scatterplot
```{r}
plot( x = prestige$education, y = prestige$income, xlab = "Education", ylab = "Income", pch = 20)
```

It looks like there is not a clear linear correlation between education and income.

#Analyse the linear relationships in the prestige data set.
```{r}
cor.test( ~ income + education,
         data=prestige,
         method = "pearson",
         conf.level = 0.95)
```

#Splitting the data set
```{r}
set.seed(123)
split <- sample.split(prestige,SplitRatio = 0.75)
train <- subset(prestige,split==TRUE)
test <- subset(prestige,split==FALSE)
```

#Creating a linear regression model
```{r}
Income_Education_LM <- lm(data = train, income ~ education)
summary(Income_Education_LM)
```

#Plotting Education against Income linear regression model
```{r}
plot(train$education, train$income, pch = 16)
curve(predict(Income_Education_LM, data.frame(education=x)), lwd = 2,  add=T)
```

#Predict the Income based on the Education values and plot the results
```{r}
test$predicted.income <- predict(Income_Education_LM, test)
predicted_plot <- test %>%
  ggplot(aes(income, predicted.income)) +
  geom_point(alpha=0.5) +
  stat_smooth(aes(colour = "black"), method = "lm") + 
  xlab("Actual value of income") +
  ylab("Predicted value of income") +
  theme_bw()
predicted_plot
```


#Calculating RMSE 
```{r}
error <- test$income-test$predicted.income
l_rmse <- sqrt(mean(error^2))
l_rmse
```

#Inspect linear model
```{r}
summary(Income_Education_LM)
```

#Look at the original scatterplot and decide which non-liner function would fit the data
The scatterplot looks to have an exponential increase after 12 therefor an exponential function would fit this dataset.

#Eyeball an exponential line of best fit on to the scatter graph
```{r}
plot( x = prestige$education, y = prestige$income, xlab = "Education", ylab = "Income", pch = 20)
curve(exp(0.6*x)+4000, lwd = 3, add = T)
```

#Plot a non-linear model using 'NLS'
```{r}
nonLinearModel <- nls(income ~ exp(b*education) + c, prestige, start = list(b = 1, c = 0))
plot(train$education, train$income, pch = 16, xlab = "Education", ylab = "Income")
curve(predict(nonLinearModel, data.frame(education=x)), lwd = 3, add=T)
```

#Look at some summary statistics for the exponential model
```{r}
summary(nonLinearModel)
```

# Plot predicted vs actual income
```{r}
test$predicted.nonLinear <- predict(nonLinearModel, test)
pl1 <- test %>%
  ggplot(aes(income,predicted.nonLinear)) +
  geom_point(alpha=0.5) +
  stat_smooth() +
  xlab('Actual value of income') +
  ylab('Predicted value of income') +
  theme_bw()
pl1
```

#Calculate the RMSE 
```{r}
error <- test$income-test$predicted.nonLinear
nl_rmse <- sqrt(mean(error^2))
nl_rmse
```

#Compare RMSE of the linear vs the non-linear model
```{r}
cat("linear model =", l_rmse,"\nnon-linear model =", nl_rmse)
```

