---
title: "Titanic Project in R"
output: html_notebook
---
# 1. Overview
Studying the survival rates from the titanic

# 2. Set up WD and import required libraries
```{r}
#Setting the working directory
setwd("C:/Users/Ashley Rainbow/Desktop/Data Fellowship/data-fellowship/R Coding")

#Load Readr
library(readr)
```

# 3. Import required files
```{r}
#Loading the train data set
train <- read_csv("train.csv")
View(train)

#Loading the test data set
test <- read_csv("test.csv")
View(test)
```

# 4. Inspect the newly imported files
```{r}
#View the class of train data set
class(train)

#View the structure of the train data set
str(train)
```

# 5. From inspecting the files I can see that the class of Pclass, Survived and Name is incorrect.
```{r}
#Set train$Pclass and train$Survived as a factor 
train$Pclass <- as.factor(train$Pclass)
train$Survived <- as.factor(train$Survived)

#Set train$Name as a character
train$Name <- as.character(train$Name)
```

# 6. Test to check for any duplicate data.
```{r}
#Check for duplicate data
print("There should be 891 unique values.")
length(unique(train$PassengerId))
```

# 7. Now take a look at the test data.
```{r}
#View the structure of the test data set
str(test)
```

# 8. You can see, like step 5, you can see that the class of Pclass and Name is set incorrectly
```{r}
#Set test$Pclass as factor
test$Pclass <- as.factor(test$Pclass) 

#Set test$Name as factor
test$Name <- as.character(test$Name)
```

# 9. Test to check for any duplicate data.
```{r}
#Check for duplicate data
print("There should be 418 unique values.")
length(unique(test$PassengerId))
```

# 10. Store survival rates to train_count_died and train_count_died. Then create proportion table. 
```{r}
#Explore the data
table(train$Survived)

#Store the values
train_survived_table <- table(train$Survived)

#Storing the count of survived and died
train_count_died <- c(train_survived_table[1])
train_count_survived <- c(train_survived_table[2])

#Creating proportions table
proportional_survived <- prop.table(table(train$Survived))
```

# 11. Plot Sex vs Survival.
```{r}
#Loading ggplot2
library(ggplot2)

#Plotting sex and survival
ggplot(train, aes(x = Sex, fill = factor(Survived))) +
  geom_bar(width = 0.5) +
  xlab("Sex") +
  ylab("Total count") +
  labs(fill = "Survived")
```

# 12. Create model to predict survival.
```{r}
#Adding a model prediction column
train$ModelPrediction <- as.factor("TBD")
head(train)

#***[this code is inneficient]***
#model_prediction <- rep(c("TBD"), each = 891)
#train <- cbind(train, model_prediction)
#colnames(train)[which(names(train) == "model_prediction")] <- "ModelPrediction"
#head(train)
#***resume normal coding***

#*******************************************************************************
class(train$ModelPrediction)
#Make all values in train$ModelPrediction "0"
train$ModelPrediction <- 0
head(train)
factor(train$ModelPrediction, levels = c(0, 1))
head(train)
#*******************************************************************************
```

# 13. Test different conditions
```{r}
#Finding the percentage accuracy of prediction when all passengers do not survive
no_survival <- sum(train$ModelPrediction == train$Survived)/nrow(train)

#Setting male prediction as 0
train$ModelPrediction[train$Sex == "male"] <- 0

#Setting female prediction as 1
train$ModelPrediction[train$Sex == "female"] <- 1

head(train)

#Finding the percentage accuracy of the prediction with females surviving
all_females_survive <- sum(train$ModelPrediction == train$Survived)/nrow(train)

#Creating a table for train classes
train_passenger_class <- table(train$Pclass)

#Assigning class counts to variables
train_count_first_class <- c(train_passenger_class[1])
train_count_second_class <- c(train_passenger_class[2])
train_count_third_class <- c(train_passenger_class[3])

#Creating proportions table for passenger class
proportional_class <- prop.table(table(train$Pclass, train$Survived))
proportional_class

#Plotting gender survival rates including class.
ggplot(train, aes(x = Sex, fill = Survived)) +
  geom_bar(width = 0.5) +
  facet_wrap(~Pclass) +
  ggtitle("Pclass") +
  xlab("Sex") +
  ylab("Total count") +
  labs(fill = "Survived")

#Setting train$Modelprediction as 0
train$ModelPrediction <- 0
head(train)

#Setting train$Modelprediction as 1 for females in first and second class
train$ModelPrediction[train$Sex == "female"] <- 1
train$ModelPrediction[train$Pclass == "3"] <- 0
head(train)

#Finding the percentage accuracy of prediction when 1st and 2nd class females survived
first_second_class_females_accuracy <- sum(train$ModelPrediction == train$Survived)/nrow(train)
paste("The first model returned an accuracy of", all_females_survive, "the second model returned an accuracy of", first_second_class_females_accuracy)

#Plotting Pclass, Sex, Age and Survival
ggplot(train, aes(x = Age, fill = Survived)) +
  facet_wrap(~Sex + Pclass) +
  geom_bar(width = 10) +
  xlab("Age") +
  ylab("Total count") 

#Resetting the train$ModelPrediction to 0
train$ModelPrediction <- 0

#Setting train$Modelprediction as 1 for females in first class, second class and for U18's 
train$ModelPrediction[train$Sex == "female"] <- 1
train$ModelPrediction[train$Pclass == "3"] <- 0
train$ModelPrediction[(train$Pclass == "1" | train$Pclass == "2") & train$Sex == "male" & train$Age < 18] <- 1

##Finding the percentage accuracy of prediction when 1st and 2nd class females survived
upper_class_female_and_u18_accuracy <- sum(train$ModelPrediction == train$Survived)/nrow(train)
paste("The first model returned an accuracy of", all_females_survive, "the second model returned an accuracy of", upper_class_female_and_u18_accuracy)
```

# 14. Predict and create export for test data set.
```{r}
#Creating submissions df
submission <- data.frame(test$PassengerId)

#Renaming the submission$test.PassengerId to PassengerId
names(submission)[names(submission) == "test.PassengerId"] <- "PassengerId"
View(submission)

#Setting survival rates to 0
submission$Survived <- as.integer(0)
factor(submission$Survived, levels = c(0, 1))

#Setting prediction values
submission$Survived[test$Sex == "female"] <- 1
submission$Survived[test$Pclass == "3"] <- 0
submission$Survived[(test$Pclass == "1" | test$Pclass == "2") & test$Sex == "male" & test$Age < 18] <- 1
View(submission)

#Export submission
write.csv(submission, file = "AshleyRainbow_Virtualstock_TitanicSubmission.csv")
```