---
title: "Regression"
output: html_notebook
---

```{r}
library(MASS)
library(corrplot)
library(ggplot2)
library(caTools)
library(dplyr)
library(plotly)
```

#Import Dataset
```{r}
boston <- Boston
```

#Explore the data using the usual summary statistics
```{r}
summary(boston)
```

```{r}
str(boston)
```

```{r}
head(boston)
```

#Cleaning the data
Remove the 'black' variable.
```{r}
boston$black <- NULL
```

#Seperate data into a test and train data sets.
```{r}
set.seed(123)
sample <- sample.int(n = nrow(boston), size = floor(.75*nrow(boston)), replace = F)
train <- boston[sample,]
test <- boston[-sample,]
```

#Finding Correlations
```{r}
cor(boston, boston$medv)
```

#Assign the correlation to a new variable
```{r}
boston_cor <- cor(boston)
boston_cor
```

#Visualise the corrolations
```{r}
corrplot(boston_cor, method = 'circle')
```

#Isolating 'medv' corrolation to other variables as it is the dependent variable that we're trying to predict.

#Build a model
##Univariate
Select one variable to predict 'medv' and use 'lm' to build a linear model of our data.
```{r}
univariate_model <- lm(medv ~ rm, data = train)
summary(univariate_model)
```

##Plot the model
```{r}
plot(univariate_model)
```

#Plot the residuals
```{r}
res <- residuals(univariate_model)
res <- as.data.frame(res)
ggplot(res, aes(res)) + geom_histogram(fill = 'blue', alpha = 0.5)
```

#Create a variable to capture predicted prices.
```{r}
test$predcited.medv <- predict(univariate_model, test)
```

```{r}
pl1<-test %>%
  ggplot(aes(medv, predicted.medv)) +
  geom_point(alpha = 0.5) +
  stat_smooth(method = "lm") +
  xlab('Actual value of medv') +
  ylab('Predicted value of medv') +
  theme_bw()
ggplotly(pl1)
```

#Use Mean Squared Error to check whether the model fits the data.

```{r}
error <- test$medv - test$predicted.medv
rmse <- sqrt(mean(error^2))
rmse
```

