---
title: "Data Fellowship Hackathon"
output: html_notebook
---
#Set WD and import libraries
```{r}
setwd("C:/Users/Ashley Rainbow/Desktop/Data Fellowship/data-fellowship/Hackathon 1")
library(tidyverse)
library(ggplot2)
library(googleVis)
```


#Import data
athlete_events contains the data regarding to the medals won. noc_regions contains the country names against their 'noc'. 
```{r}
olympics <- read.csv("athlete_events.csv")
NOC <- read.csv("noc_regions.csv")
```

#Assign values to meadals
This will allow me to sum the score for each country.
```{r}
olympics$Value[olympics$Medal == "Gold"] <- 3
olympics$Value[olympics$Medal == "Silver"] <- 2
olympics$Value[olympics$Medal == "Bronze"] <- 1
olympics$Value[is.na(olympics$Medal)] <- 0
```

#Structuring the data
Adding a country column to olympics data based on the 'NOC' field. 
Due to some countries from past olympics no longer existing we only filter for events after the year 2000. 
```{r}
olympics <- merge(olympics,NOC,by="NOC")
modern_olympics <- subset(olympics, olympics$Year >= 2000)
```

#Country Scores
Grouping on the country name and summing the 'Value' column allows me to calculate the score for each country. 
```{r}
country_collapse <- modern_olympics %>% group_by(region) %>% summarise(Total = sum(Value))
```

#Tidying the data
Due to some mismatches between the country column and Google's naming convention we have to adjust some of the names.
There is also an na field in the data so this must be removed.
```{r}
country_collapse <- na.exclude(country_collapse)
country_collapse$region <- as.character(country_collapse$region) 
country_collapse$region[country_collapse$region == 'USA'] <- 'United States'
country_collapse$region[country_collapse$region == 'UK'] <- 'United Kingdom'
country_collapse$region[country_collapse$region == 'Boliva'] <- "Bolivia"
country_collapse$region[country_collapse$region == 'Republic of Congo'] <- 'CG'
country_collapse$region[country_collapse$region == 'Democratic Republic of the Congo'] <- 'CD'
country_collapse$region[country_collapse$region == 'South Sudan'] <- 'SS'
```

#Plotting the Map
The map opens into a new web browser tab.
```{r}
map <- gvisGeoChart(country_collapse, locationvar = 'region', colorvar = 'Total', hovervar = 'Total', options = list(width=1000, height=700))
plot(map)
```
