---
title: "Predicting House Prices"
output: html_notebook
---

Grab your libraries. You'll need the following for this tutorial.
```{r}
library(MASS)
library(corrplot)
library(ggplot2)
library(caTools)
library(dplyr)
library(plotly)
```

# Import Boston dataset
'Boston' is included in the 'MASS' package
```{r}
boston <- Boston
```

# Explore the data
```{r}
summary(boston)
```

```{r}
str(boston)
```

```{r}
head(boston)
```

*definitions for the data fields:* https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/Boston.html

`MEDV` is the price variable. That's our dependent variable.

#Cleaning the data
Remove the 'black' variable.
```{r}
boston$black <- NULL
```

#Split the data between test and train

```{r}
set.seed(123) 
sample <- sample.int(n = nrow(boston), size = floor(.75*nrow(boston)), replace = F)
train <- boston[sample, ] 
test  <- boston[-sample, ]
```

#Finding correlations

```{r}
cor(boston,boston$medv)
```

Assign the correlations to a variable
```{r}
boston_cor <- cor(boston)
boston_cor
```

#Visualise these corrolations
```{r}
corrplot(boston_cor, method = "circle")
```


# Build a univariate model
`medv` and `lm` will be used to build a linear model of our data.
```{r}
univariate_model <- lm(medv ~ rm, data=train)
summary(univariate_model)
```

#Plot the new model.
```{r}
plot(univariate_model)
```

Use `abline` to plot your new model against the two variables you've chosen (one will be `medv`).

Plot the residuals.
```{r}
# Use residuals to grab the residuals of the univariate model and pop it into a variable called `res`
res <- residuals(univariate_model)
# Convert residuals to a data frame so that it can be used elsewhere
res <- as.data.frame(res)
# Use ggplot to create a histogram
ggplot(res,aes(res)) +  geom_histogram(fill='blue',alpha=0.5)
```

#Capture predicted prices.
```{r}
test$predicted.medv <- predict(univariate_model,test)
```

#Plot it
```{r}
pl1 <- test %>% 
  ggplot(aes(medv,predicted.medv)) +
  geom_point(alpha=0.5) + 
  stat_smooth(method = "lm") +
  xlab('Actual value of medv') +
  ylab('Predicted value of medv') +
  theme_bw()
ggplotly(pl1)
```

#Judge whether or not the model fits the data.
```{r}
error <- test$medv-test$predicted.medv
rmse <- sqrt(mean(error^2))
rmse
```

#Multivariate model
Repeat the previous steps with multiple variables
```{r}
multivariate_model <- lm(medv ~ .,data=train)
summary(multivariate_model)
```

#Plot those results.
```{r}
plot(multivariate_model)
```

#Check the residuals.
```{r}
res <- residuals(multivariate_model)
# Convert residuals to a DataFrame 
res <- as.data.frame(res)
ggplot(res,aes(res)) +  geom_histogram(fill='blue',alpha=0.5)
```

#Plot it.
```{r}
test$predicted.medv <- predict(multivariate_model,test)
pl2 <- test %>% 
  ggplot(aes(medv,predicted.medv)) +
  geom_point(alpha=0.5) + 
  stat_smooth(aes(colour='black'), method = "lm") +
  xlab('Actual value of medv') +
  ylab('Predicted value of medv') +
  theme_bw()
ggplotly(pl2)
```

#Check the error
```{r}
error <- test$medv-test$predicted.medv
rmse <- sqrt(mean(error^2))
rmse
```
