---
title: "Black Friday Clustering Using All Methods"
output: html_notebook
---
# Problem Statement.
We are unsure on how the suppliers performed over Black Friday for Retailer 1. It would be useful to see how they performed to understand how much Retailer 1 is benefitting from these suppliers.

# Suggested solution.
Use K-means clustering, DBSCAN and APcluster to group together suppliers based on the quantity of sales vs the average value of their sales.

# 1. Overview

This notebook uses k-means clustering, Affinity Propagation and DBscan to group suppliers for Retailer 1 based on the quantity of sales vs the average value of their sales.

# 2. Load libraries

Tidyverse, dbscan and apcluster will cover everything I need to do.
```{r}
#install.packages("tidyverse")
#install.packages("dbscan")
#install.packages("apcluster")
library(tidyverse)
library(dbscan)
library(apcluster)
```

# 3. Load data

Import the CSV file for Retailer 1 into R, and use 'head' to preview the new data.
```{r}
data <- read.csv('BF Clusterings.csv')
head(data)
```

# 4. Explore and transform the data

Looking at the data it is clear to see that the column headers can be renamed to make it easier for 3rd parties to understand the data.
```{r}
# The column headers are confusing due to the symbols I will rename them to replace the '.' with a space.
names(data) <- c("Supplier", "Average Order Cost (£)", "Count of Orders")

# I will check the the changes have worked as expected.
head(data)
```

# 5. Visualise the data

Now that my data is in the format I want, I can visualise this data in a scatter plot. This will give me a good chance to eyeball the data and get an idea of what I might expect to see from the final result.
```{r}
qplot(x = data$`Average Order Cost (£)`, y = data$`Count of Orders`, label = data$Supplier, geom = "text")
```


# K-MEANS
# 1. Fine tuning k-means

Use WSS to determine the optimal amount of clusters.
```{r}
kClustersMax <- 10

wss <- sapply(1:kClustersMax,
        function(k){kmeans(data[,2:3], k)$tot.withinss})

plot(1:kClustersMax, wss,
       type="b", pch = 19, frame = FALSE,
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")
```

# 2. Clustering

R has a built in `kmeans` function that I can use to create my clusters. [This documentation](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/kmeans.html) explains how to use it.
```{r}
# Use the wss graph from the previous step to decide the optimum number of clusters.
kClusters <- 5

# Use the kmeans function and store the result in a variable
kMeansResult <- kmeans(data[,2:3], centers = kClusters)

# Create a new column in your dataset to show which cluster each ward belongs to
data$cluster <- kMeansResult$cluster

# Make sure your new cluster column is a factor, not an integer
data$cluster <- as.factor(data$cluster)

# Export clustered suppliers as CSV
write.csv(data, file = "Clustered BF.csv")

head(data)
```

# 3. Visualise the clusters

Use qplot again to visualise the clustered data on the same scatterplot.
```{r}
qplot(x = data$`Average Order Cost (£)`, y = data$`Count of Orders`, label = data$Supplier, geom = "text", color = data$cluster)
```


# AFFINITY PROPAGATION
# 1. Reset the data

```{r}
data <- read.csv('BF Clusterings.csv')
names(data) <- c("Supplier", "Average Order Cost (£)", "Count of Orders")
```

# 2. Use APCLUSTER to cluster the data.

AP clustering can be used to group the Black Friday data.
Useful resources:
[How affinity propagation works](https://medium.com/@aneesha/using-affinity-propagation-to-find-the-number-of-clusters-in-a-dataset-52f5dd3b0760)

Looking at the WSS method from the K-means clustering I can see that the optimum number of clusters is ~5
```{r}
clustered <- apclusterK(negDistMat(r=2), data, K = 5)
show(clustered)
```

# 3. Plot out the clusters.

Now that I have clustered my data I can visualise this using the plot function.
```{r}
plot(clustered, data[,2:3], xlab = "Average Value of Sale (£)", ylab = "Count of Sales")
```


#DBSCAN
# 1. Reset the data

```{r}
data <- read.csv('BF Clusterings.csv')
names(data) <- c("Supplier", "Average Order Cost (£)", "Count of Orders")
```

# 2. Clustering using DBSCAN and LOF (Local Outlier Factor)

I will use DBSCAN to cluster the Black Friday data. 
Useful resources:
[How DBSCAN works](https://towardsdatascience.com/how-dbscan-works-and-why-should-i-use-it-443b4a191c80)
[How LOF works](https://towardsdatascience.com/local-outlier-factor-for-anomaly-detection-cc0c770d2ebe)
[Source for DBSCAN and LOF code](https://github.com/mhahsler/dbscan)

I will start by using DBSCAN to see the cluster in this data set.
```{r}
x <- as.matrix(data[, 2:3])
db <- dbscan(x, eps = 50, minPts = 1)
db
```

Now I can use a scatter plot to visualise these clusters.
```{r}
plot(x, col = db$cluster + 1L)
```
I can also visualise this using 'pairs' to clearly show the 2 variables
```{r}
pairs(x, col = db$cluster + 1L)
```

LOF can be used to show outliers in this data set. (Probably more useful in larger data sets)
```{r}
lof <- lof(x, k = 8)
plot(x, cex = lof)
```

# 3. Use HDBSCAN to heirachically cluster the data set.
First, cluster the groups using the hdbscan function.
```{r}
hdb <- hdbscan(x, minPts = 2)
hdb
```

Now I can visualise this using plot.
```{r}
plot(hdb, show_flat = T, gradient = c("yellow", "orange", "red", "blue"))
```

#Summary
From this you can begin to look at the KMeans clusting, AP clustering and the DBSCAN scatter plots. The suppliers in the bottom left hand corner can be questioned whether they contribute value to the retailers range as they sell so few products at a low value. The suppliers in the low center and the bottom right can be looked into how they can push the suppliers products better, this is as the retailer can profit more off of selling greater quantites due to their higher value products. The supliers in the top and mid left are worth trying to maintain as they sell at such high quantities it is a steady income for the retailer.