---
title: "South Oxfordshire Hoodmaps with Code"
output: html_notebook
---

# 1. Overview
This notebook uses k-means clustering to group areas in South Oxfordshire.

# 2. Load libraries

The Tidyverse collection of libraries includes all of the packages required to complete this analysis.

This is also a good point to set your working directory to the folder you'll use to save your data.

```{r}
install.packages("tidyverse")

library(tidyverse)
```

# 3. Load data
Import the CSV file for South Oxfordshire Statisticsinto R.
```{r}
data <- read.csv('Soxon Statistics.csv')
```

# 3. Explore and transform your data

Take a look at the data you've just imported to make sure everything loaded properly.

After a quick glance, I noticed there were a few changes I wanted to make.

```{r}
# The column headers are lengthy. I want to change each of them to be shorter and clearer.
names(data) <- c("Area", "Average Age", "Unployment Rate (%)")

# Now I can take a look at my data and see how it has changed.
head(data)
```

# 4. Visualise your data

Now that my data is in the format I want, I can take a quick look at it using qplot.

```{r}
qplot(x = data$`Average Age`, y = data$`Unployment Rate (%)`, label = data$Area, geom = "text")
```

# 5. Clustering

R has a built in `kmeans` function that I can use to create my clusters. [This documentation](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/kmeans.html) explains how to use it.

```{r}
# Let's guess that our dataset has 3 clusters.
kClusters <- 3

# Let's use the kmeans function and store the result in a variable
kMeansResult <- kmeans(data[,2:3], centers = kClusters)

# The code above can also be written as kMeansResult <- kmeans(data[,2:3], centers = kClusters)

# Create a new column in your dataset to show which cluster each ward belongs to
data$cluster <- kMeansResult$cluster

# Make sure your new cluster column is a factor, not an integer
data$cluster <- as.factor(data$cluster)

View(data)
```

# 6. Visualise your clusters

Use qplot again to

```{r}
qplot(x = data$`Average Age`, y = data$`Unployment Rate (%)`, label = data$Area, geom = "text", color = data$cluster)
```

# Fine tuning k-means

Use wss to determine the optimal amount of clusters.
```{r}
kClustersMax <- 10

wss <- sapply(1:kClustersMax,
        function(k){kmeans(data[,2:3], k)$tot.withinss})

plot(1:kClustersMax, wss,
       type="b", pch = 19, frame = FALSE,
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")
```