---
title: "Anscombe Exercise"
output: html_notebook
---
# Set up
```{r}
setwd("C:/Users/Ashley Rainbow/Desktop/Data Fellowship/data-fellowship/Data Visualisation/Data Viz Exercise")
install.packages("ggplot2")
library(ggplot2)
```

# Import Data
```{r}
data("anscombe")
```

# Create a summary for anscombes data
```{r}
summary(anscombe)
```

# Plot anscombes graphs
```{r}
plot(anscombe$x1, anscombe$y1, xlim = c(0, 20), ylim = c(0,14))
abline(lm(anscombe$y1~anscombe$x1), col="red")
```
You can see here that the values have a large distribution but as x1 increases y1 increases with it.

```{r}
plot(anscombe$x2, anscombe$y2, xlim = c(0, 20), ylim = c(0,14))
abline(lm(anscombe$y2~anscombe$x2), col="red")
```
Here you can see a more quadratic graph where if x2 contiued to aproximately 17 y2 would return to 3, but as x2 stops at 14 it has the same trend as x1/y1.

```{r}
plot(anscombe$x3, anscombe$y3, xlim = c(0, 20), ylim = c(0,14))
abline(lm(anscombe$y3~anscombe$x3), col="red")
```
Due to the anomaly at x=13 the trend line matches x1/y1 and x2/y2.

```{r}
plot(anscombe$x4, anscombe$y4, xlim = c(0, 20), ylim = c(0,14))
abline(lm(anscombe$y4~anscombe$x4), col="red")
```
This, much like x3/y3, has an anomaly that drastically changes the trend line. Without the annomaly at x = 19 the trend line would be x = 8
