---
title: "Black Friday Clustering Using Affinity Propagation (AP)"
output: html_notebook
---
# Problem Statement.
We are unsure on how the suppliers performed over Black Friday for Retailer 1. It would be useful to see how they performed to understand how much Retailer 1 is benefitting from these suppliers.

# Suggested solution.
Use AP clustering to group together suppliers based on the quantity of sales vs the average value of their sales.

# 1. Overview

This notebook uses Affinity Propagation clustering to group suppliers for Retailer 1 based on the quantity of sales vs the average value of their sales.

# 2. Load libraries

I can use Tidyverse and apcluster to complete this task.

```{r}
#install.packages("tidyverse")
#install.packages("apcluster")
library(tidyverse)
library(apcluster)
```

# 3. Load data

Import the CSV file for Retailer 1 into R, and use 'head' to preview the new data.

```{r}
data <- read.csv('BF Clusterings.csv')
head(data)
```

# 4. Explore and transform the data

Looking at the data it is clear to see that the column headers can be renamed to make it easier for 3rd parties to understand the data.

```{r}
# The column headers are confusing due to the symbols I will rename them to replace the '.' with a space.
names(data) <- c("Supplier", "Average Order Cost (£)", "Count of Orders")

# I will check the the changes have worked as expected.
head(data)
```

# 5. Visualise the data

Now that my data is in the format I want, I can visualise this using a scatter plot. This will give me a good chance to eyeball the data and get an idea of what I might expect to see from the final result.

```{r}
qplot(x = data$`Average Order Cost (£)`, y = data$`Count of Orders`, label = data$Supplier, geom = "text")
```

# 6. Use APCLUSTERS to cluster the data.

After seeing the data you can get an idea of how many clusters there may be. In this case I will say 4 other methods can be used to get a more accurate idea of the optimum number of clusters.

```{r}
clustered <- apclusterK(negDistMat(r=2), data, K = 4)
show(clustered)
```

# 8. Plot out the clusters.

Now that I have clustered my data I can visualise this using the plot function.
```{r}
plot(clustered, data[,2:3], xlab = "Average Value of Sale (£)", ylab = "Count of Sales")
```