---
title: "Association Rules Project"
output:
  html_notebook: default
  word_document: default
---
# Problem Statement
We want to assist the retailer and help them to sell more products using an Amazon style "People who bought this also bought:" system.

# Suggested Solution
Using Association rules I can find common itemsets that are bought together to help the retailer start with the most popular item sets.

# 1.Load libraries
Here I will load in dplyr, arules and arulesViz. Dplyr is a very good tool for data transformation this will help me get the data into a suitable state to analyse the data. Arules is used to analyse data to find common itemsets and association rules. ArulesViz enables visualisation of the results we gain from the arules package.
```{r}
library(dplyr)
library(arules)
library(arulesViz)
```

# 2.Set the working directory
```{r}
setwd("C:/Users/Ashley Rainbow/Desktop/Data Fellowship/data-fellowship/Association Rules/Association Rules Project/Association Rules project 2")
```

# 3.Import the data
```{r}
orders <- read.csv('Peak Association Rules.csv')
```

# 4.Check the data
This allows me to see if the data is in a suitable structure.
```{r}
head(orders)
```
```{r}
str(orders)
```

# 7.Group products by order number
This allows me to show products bought in an order as one collective 'basket'. 
```{r}
order_baskets = orders %>% 
  group_by(Order.Number) %>% summarise(order_basket = as.vector(list(Item.Name)))
```

# 8.View a sample order_basket
Check to see if dataframe is set up correctly (order basket is set up as a list of factors).
```{r}
head(order_baskets)
```

# 9.Convert to sparse matrix
To view itemsets and association rules the dataframe needs to be converted to a sparse matrix.
```{r}
sparse_data <- as(order_baskets$order_basket, "transactions")
```

# 10.Inspect the sparse matrix
```{r}
summary(sparse_data)
inspect(sparse_data[1:20])
```

# 11.Plot the most popular products
This gives me an idea of what products sell well and what to keep an eye out for later on.
```{r}
par(mar = c(4, 5, 1, 1))
itemFrequencyPlot(sparse_data, topN = 10, type = 'absolute', cex.names = 0.5, horiz = TRUE, xlim = c(0, max(1400)), col = c("#E64097"))
```

# 12.Create a rule for the most frequent itemsets
This finds the most common data sets with a support greater than 0.0004.
```{r}
rules = apriori(data = sparse_data, parameter = list(target = "frequent itemsets",  support = 0.0004, minlen = 2), control = list(verbose = FALSE))
rules = sort(rules, by = 'support', decreasing = TRUE)
inspect(rules[1])
```

# 13.Create a dataframe base on rules data.
This gives me a dataframe of the most popular itemsets.
```{r}
order_rules_df <- DATAFRAME(sort(rules, by = "support", decreasing = FALSE))
```

# 14.Create a barplot from the orders dataframe.
I use this to show the top itemsets. This give a clear demonstration of the top itemsets and how much support they have.
```{r}
par(mar = c(5,17, 1, 1))
barplot(order_rules_df$support, main = "Order Rules", xlab = "Support", names.arg = order_rules_df$items, horiz = TRUE, las = 2, cex.names = 0.5, xlim=c(0, max = 0.001), col = c("#E64097"))
```

# 15.Finding the most common itemsets
Uses apriori to find most common itemsets with support of at least 0.0004 and confidence of greater than 0.2
```{r}
rules = apriori(data = sparse_data, parameter = list(support = 0.0004, confidence = 0.2, minlen = 2))
```

```{r}
summary(rules)
```

```{r}
inspect(rules[1:10])
```

# 16.Plotting most common sets of items being bought together.
This gives a visual representation of the most common itemsets and the relationship between those products.
```{r}
plot(rules, method = 'scatterplot', measure = c("support", "confidence"), shading = "lift", jitter = 1)
```

```{r}
plot(rules[1:10], method = 'graph', control = list(cex = 0.65))
```

# 17.Viewing rule sorted by sorted by confidence.
This gives me a different perspective on the item sets seeing the top 10 most confident products. 
```{r}
inspect(sort(rules, by = 'confidence', decreasing = TRUE)[1:10])
```

# 18.Viewing rule sorted by sorted by support.
As above this shows you the top 10 most popular item sets sorted by support.
```{r}
inspect(sort(rules, by = 'support', decreasing = TRUE)[1:10])
```

# 19.Viewing the rule sorted by lift.
As above this shows you the top 10 most popular item sets sorted by lift.
```{r}
inspect(sort(rules, by = 'lift', decreasing = TRUE)[1:10])
```

# 20.Finding the most common set of items bought with higher support but lower confidence.
```{r}
rules = apriori(data = sparse_data, parameter = list(support = 0.00055 , confidence = 0.1 , minlen = 2, maxlen = 3))
```

```{r}
summary(rules)
```

```{r}
inspect(rules)
```

# 21.Plotting most common sets of items being bought together with low confidence but high support.
This allows me to visualise the new item sets and compare to previous sets.
```{r}
plot(rules, method = 'scatterplot', measure = c("support", "confidence"), shading = "lift", jitter = 1)
```

```{r}
plot(rules, method = "graph", control = list (cex = 0.65))
```

# 22.Creating Cream Toaster subset
This step is used to view the products commonly bought with a cream toaster.
```{r}
rules = apriori(data = sparse_data, parameter = list(support = 0.00003, confidence = 0.5, minlen = 2))
```

```{r}
cream_toaster_rule = sort(
  subset(rules, subset = rhs %in% 'Retro ST19020CN 4-Slice Toaster - Cream'),
  by = 'support', decreasing = T
)
inspect(cream_toaster_rule)
```

```{r}
lift_cream_toaster = sort(cream_toaster_rule, by = 'lift', decreasing = TRUE)
paste("Sorted by lift. ----------------------------------")
inspect(lift_cream_toaster)
```

```{r}
confidence_cream_toaster = sort(cream_toaster_rule, by = 'confidence', decreasing = TRUE)
paste("Sorted by confidence. ----------------------------------")
inspect(confidence_cream_toaster)
```

This is used to visualise the products relationships with the cream toaster.
```{r}
plot(cream_toaster_rule, method = 'graph', control = list(cex = 0.7))
```

# 23.Generate summary reports for summary and lift rules.
```{r}
paste("Rules Report")
paste("Support")
summary(rules@quality$support)
paste("Lift")
summary(rules@quality$lift)
```

```{r}
paste("Cream Toaster Rules Report")
paste("Support")
summary(rules@quality$support)
paste("Lift")
summary(rules@quality$lift)
```

# 24.Comparing Cream Toaster rule results to the other rules.
```{r}
support_list <- rules@quality$support
mean_cream_toaster_support <- mean(cream_toaster_rule@quality$support)
lower_than_mean_rules <- support_list[support_list>c(mean_cream_toaster_support)]
support_rank <- length(lower_than_mean_rules)
support_rank
```
```{r}
paste("On average, the Cream Toaster has the", support_rank, "largest support of all the", length(rules), "rules.")
```

```{r}
confidence_list <- rules@quality$confidence
mean_cream_toaster_confidence <- mean(cream_toaster_rule@quality$confidence)
lower_than_mean_rules_confidence <- confidence_list[confidence_list>c(mean_cream_toaster_confidence)]
confidence_rank <- length(lower_than_mean_rules_confidence)
confidence_rank
```
```{r}
paste("On average, the Cream Toaster has the", confidence_rank, "largest confidence of all the", length(rules), "rules.")
```

```{r}
paste("In total", length(orders$Item.Name=="Retro ST19020CN 4-Slice Toaster - Cream"), "cream toasters were sold")
```

# Summary
As you can see above, the cream toaster is commonly bought with products of the same colour from the same brand. This suggests that selling these products as a bundle, for example a cream kitchen combination set, which would include a cream kettle and a cream bread bin could result in an increase in sales for products of this group.

>{Retro 4-Slice Toaster, Retro Jug Kettle, Retro Towel Pole, Retro Bread Bin Cream, Retro Microwave with Grill, Retro Traditional Kettle, Retro 45-litre Rectangular Bin, Solo Microwave, Retro 1-litre Canisters, Retro 2-piece Non-stick Frying Pan Set}